### 2021-07-11 submission/submission.tex

% !TeX spellcheck = en_US
\documentclass[10pt,a4paper]{article}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=blue,
	filecolor=magenta,
	urlcolor=gray
}

\usepackage{xcolor}
\usepackage{listings}

\lstdefinestyle{BashInputStyle}{
	language=bash,
	basicstyle=\small\sffamily,
	columns=fullflexible,
%	backgroundcolor=\color{yellow!20},
	linewidth=0.9\linewidth,
	xleftmargin=0.1\linewidth
}

\author{OcamlPro - Fabrice Le Fessant et Thomas Sibut-Pinote}
\title{Groth16 zkSNARK Proof Verification on Freeton : Three Use Cases}

\begin{document}
\maketitle

The present document is an official submission to the 18$^{\text{th}}$
contest\footnote{\url{https://devex.gov.freeton.org/proposal?proposalAddress=0\%3Ae6b65075478e7d412fdb0870452f30dfa8bf51272e28a3167abc5c5df6fd051d}}
of the DevEx governance: {\em Groth16 zkSNARK Proof Verification Use
  Cases}. Although we accepted earlier to be part of the jury, we
understand that by submitting this document we waive our right to vote
in the present contest.

In this submission, we propose three different use cases for Groth16
zkSNARK proofs on the FreeTON blockchain:
\begin{itemize}
\item Sudoky: Zksnarks are used to prove that the user has a solution
  to a given Sudoky problem, without providing the solutions to other
  users ;
\item Euler Project: Zksnarks are used to prove that the user has
  found the solution to an Euler problem ;
\item Pin-code reset of forgotten keys: Zksnarks are used to replace a
  pubkey when the user has lost his secret key
\end{itemize}

Finally, we finish this document by a section showing the
contributions that we did for the Free TON community while working on
this contest.

\section{A toy example: Sudoku}
\label{section_sudoku}

Let's start with our first toy use case. Suppose you want to set up a Sudoku problem for your students to solve, so that upon completion they receive some token. The issue is that once any student has solved it, the solution sits on the blockchain for all to see. All the other students can cheat by copying it, instead of doing the work by themselves, which defeats the whole purpose. One other advantage is that we don't need to implement the whole logic of the computation on the smart contract. For some size of Sudoku, this is probably worthwhile.

\section{A better example: Project Euler}
\label{section_euler}

The previous example is somewhat limited by the fact that sudokus are easily solvable problems, either by hand or with a computer; they are not a realistic use case of zk-snarks. Let's keep the idea of a decentralized classroom. the famous website Project Euler\footnote{\url{https://projecteuler.net/}} offers math and programming problems of various difficulty to its users. It uses captchas to prevent brute force attempts at guessing the answer to a problem. The answers to problems are stored statically on its servers and the answers are checked against them. Can we make Project Euler decentralized? Using the idea from Section \ref{section_sudoku}, we can prevent users from stealing others' solutions \textbf{and} from brute-forcing them, due to the cost of generating proofs.

\section{Pin code reset for forgotten keys}
\label{section_pincode}

In this use case, we use Zksnarks to allow users to replace their
public keys in multisig wallets. The typical use is when a user has
lost the associated secret key, and is unable to sign/confirm any new
transaction. With our approach, the user first creates a recovery
contract, where a passphrase is associated to his public key. If he
needs to change his public key, he can use Zksnarks to associate a new
public key in his recovery contract, and tell multisig wallets to
safely replace his pubkey by the new one.

Compared to other approaches, this use can provide several
improvements:
\begin{itemize}
\item The user does not disclose his passphrase. Most other approaches
  (based on hashes for example) give an opportunity for
  man-in-the-middle attacks ;
\item As the user does not disclose his passphrase, the pubkey can be
  replaced several times ;
\item The user does not need to create the new public key ahead of
  time. Instead, he can create it when he has lost the former one,
  decreasing the likelihood of losing the associated secret key also ;
\end{itemize}

\subsection{Technical solution}

This use case includes 3 different components:
\begin{description}
\item[PinCode Client:] it is a C++ program linked to the Blueprint
  Zksnarks library. It can be used in two different ways:
  \begin{itemize}
  \item To create the initial circuit using the passphrase. The client
    should be called as:
\begin{verbatim}
pincode-client prepare "my pass phrase"
\end{verbatim}
   This will generate a file \lstinline{verifkey.hex} that can be used
   as an argument when deploying the \lstinline{PubkeyRecovery} smart
   contract, and a file \lstinline{provkey.hex} that can be used to
   create new witnesses ;
 \item To create a witness that the passphrase is known when providing
   a new public key. Because the witness contains the new public key,
   an attacker cannot intercept the message and replace it. The client
   should be called as:
\begin{verbatim}
pincode-client prove "my pass phrase" "PUBKEY_AS_HEX_NUMBER"
\end{verbatim}
  The command expects to find the file \lstinline{provkey.hex} and
  generates a file \lstinline{proof.hex} that should be submitted to
  the \lstinline{PubkeyRecovery} smart contract together with the new
  public key.
  \end{itemize}
\item[PubkeyRecovery smart contract:] This smart contract is a TIP-3
  contract, using the initial public key to verify its address. This
  contract is very simple. It contains:
  \begin{itemize}
  \item A constructor to set the circuit that will be used to verify
    that the passphrase is known when it is used ;
  \item An external function
    \lstinline{SetNewPubkey(bytes proof, uint256 pubkey)}
    to define the new public key, while providing a
    proof of knowledge of the passphrase ;
  \item An external function
    \lstinline{RecoverPubkey(address multisigaddr)}
    to call a multisig contract to update the
    corresponding custodian. The multisig contract will verify that
    the address is correct before replacing the public key ;
  \end{itemize}
\item[Modified Multisig Wallet:] This is a standard multisig wallet,
  modified to recognize addresses of \lstinline{PubkeyRecovery} smart
  contracts. They provide two additionnal functions:
  \begin{itemize}
  \item An external function
    \lstinline{SetPubkeyRecoveryCode(TvmCell code)}
    to define the code of the \lstinline{PubkeyRecovery} smart
    contract. Because the code hash is known, anybody can call this
    function with the correct code.
  \item An external function
    \lstinline{RecoverPubkey(uint256 oldkey, uint256 newkey)}
    that can only be called by a
    \lstinline{PubkeyRecovery} smart contract.
  \end{itemize}
\end{description}

\section{Contributions to the Free TON community}

In the course of participating to this contest, we were led to refine
our in-house tools in order to more easily manipulate the Nil
Foundation forks of the Ton Virtual Machine (TVM), the TON Solidity
compiler and the TVM linker.

Our first contribution is a public Docker image containing the TONOS
SE compiled with Zksnarks support. The image is published as {\tt
  ocamlpro/nil-local-node} and is built from
\url{https://github.com/NilFoundation/tonos-se}.

Running it is simple:
\begin{verbatim}
docker run -d --name local-node -e USER_AGREEMENT=yes -p80:80 ocamlpro/nil-local-node
\end{verbatim}

Another way to use it is through our development tool {\tt ft}, a Free
TON Wallet designed for developers:
\url{https://github.com/OCamlPro/freeton_wallet}

For this contest, a set of improvements has been contributed to {\tt
  ft} to ease using it.

In particular:
\begin{itemize}
\item The interface has been improved to provide multiple levels of
  subcommands, making them easier to understand and to use ;
\item A new argument {\tt --image} is available with {\tt ft switch
  create} to specify the Docker image to use. This new argument is
  specially useful for our Docker image for NilFoundation TONOS SE.
\end{itemize}

As a consequence, a Zksnark-ready sandbox (local network) can be
installed by simply running:

\begin{lstlisting}[style=BashInputStyle]
ft switch create sandbox --image ocamlpro/nil-local-node
\end{lstlisting}

Now all commands from the \lstinline|ft| documentation will work, in
particular it will be easy to create accounts, deploy contracts, and
call them as seen in
\url{https://ocamlpro.github.io/freeton_wallet/sphinx/use-cases.html}.

The Docker image of \lstinline|ft| has been updated to use
NilFoundation tools
(\lstinline|solc|, \lstinline|tonos-cli|, \lstinline|tvm_linker|), it is
the easiest way to use \lstinline|ft| if you don't want to build
it. See the installation instruction at
(\url{https://ocamlpro.github.io/freeton_wallet/sphinx/install.html#using-docker}).

\end{document}